Dish: Distribution-oblivious Shell Scripting

### 2019-11-22 pldi/draft.tex
Clean up and pass on 4

Some commands are easy to parallelize, because they maintain trivial state and are commutative---\eg \ttt{wc}, which simply maintains a counter.
Others, such as \ttt{sha1sum}, maintain more complex state that is not applied commutatively.

Often these commands do not operate in an online fashion, but need to block until the end of a stream---a typical example of this is \ttt{sort}, which cannot start before the last line of input has been consumed.
Non-streaming constraints affect task parallelism, but not data parallelism:
  \ttt{sort} can be parallelized significantly using divide-and-conquer techniques.

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce\kk{cite} and Apache Spark\kk{cite}. \todo{Unfortunately, the
  dataflow graph model of most of these systems is not expressive
  enough to represent data parallelism, so the distributed generations
  that they generate, can not be represented in their dataflow model.}
\kk{This last sentence is extremely important and we should make it
  shine.}

@@

Before describing the different types of parallelization
optimizations, we formalize our intuition about stateless and pure
commands.
\nv{TODOs
  (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me);
  (ii) try to add a few paragraphs / subsections on parallelization optimizations that make use of these formalisms (to show their value);
  (iii) try to add a proof \emph{sketch} pointing to these formalisms (to show their value), and we can point to the TR for the longer proof.
}
 Recall the example \ref{} from section Categories
\ref{}. This example shows the stateless command x being applied to
...\kk{Complete this after we have an example}. Formally, a commmand
$f$ is stateless if it is ``linear'' with respect to concatenation,
i.e. the following equation holds:

@@

Based on these equations, we can define a maximal node split operation
on a stateless node $v$ with $n$ input ordered edges and an output
edge to node $v'$, to replace $v$ with $n$ new nodes, giving each one
of them one of the incoming edges of $v$, and routing them in order to
point to $v'$. The maximal node split operation can similarly extended
for pure nodes.

\kk{If there is time I can work out a formal definition and a proof
  sketch why this transformation preserves the output of the dataflow
  graph}.

@@

\kk{Remember to mention the assumptions that need to hold for the
  graph transformations to be valid in the Command categories
  section. Commands must be deterministic, they must not do any other
  side effect (such as writing to other files, sending signals,
  etc). However, these assumptions must already be checked when the
  developer designates the categories.}

>>>>

Some commands are easy to parallelize, because they maintain trivial state and are commutative---\eg \ttt{wc}, which simply maintains a counter.
Others, such as \ttt{sha1sum}, maintain more complex state that has to be updated in a sequential fashion.

Often these commands do not operate in an online fashion, but need to block until the end of a stream---a typical example of this is \ttt{sort}, which cannot start before the last line of input has been consumed.
Non-streaming constraints affect task parallelism, but not data parallelism:
  \ttt{sort} can be parallelized significantly using divide-and-conquer techniques, that is implementing it as a map and a reduce function.

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce\kk{cite} and Apache Spark\kk{cite}.Unfortunately, the
dataflow graph model of most of these systems is not expressive enough
to represent data parallelism, so the distributed implementations that
they perform do not preserve the semantics of their dataflow model.

@@

Before describing the different types of parallelization
optimizations, we formalize our intuition about stateless and pure
commands.
\nv{TODOs
  (i) highlight earlier the antithesis between the formalisms in this section and the informalities of the previous (me);
  (ii) try to add a proof \emph{sketch} pointing to these formalisms (to show their value), and we can point to the TR for the longer proof.
}
 Recall the example \ref{} from section Categories
\ref{}. This example shows the stateless command x being applied to
...\kk{Complete this after we have an example}. Formally, a commmand
$f$ is stateless if it is ``linear'' with respect to concatenation,
i.e. the following equation holds:

@@

\heading{Parallelization Transformations}
%
Based on these equations, we can define a maximal \emph{node split
  operation} on a stateless node $v$ with $n$ input ordered edges and
an output edge to node $v'$, to replace $v$ with $n$ new nodes, giving
each one of them one of the incoming edges of $v$, and routing them in
order to point to $v'$. Since each incoming edge represents a stream
of data $x_i : D^*$, and the only behaviour of a dataflow graph is its
output, this optimization $ v(x_1.x_2...x_n) \Rightarrow
v(x_1).v(x_2)...v(x_n)$ can be shown to preserve the behaviour of the
graph. Furthermore, it can be straightforwardly extended to pure
commands that can be implemented by a map-reduce pair $(m, r)$ as $
v(x_1.x_2...x_n) \Rightarrow r(m(x_1).m(x_2)...m(x_n))$.

\tr{Remember to mention the assumptions that need to hold for the
  graph transformations to be valid in the Command categories
  section. Commands must be deterministic, they must not do any other
  side effect (such as writing to other files, sending signals,
  etc). However, these assumptions must already be checked when the
  developer designates the categories.}

\tr{If there is time I can work out a formal definition and a proof
  sketch why this transformation preserves the output of the dataflow
  graph.}

@@
